---
title: "Random uncertainty"
author: "Thomas Wutzler"
# output:
#   ioslides_presentation:
#     css: slidesTw.css
#     wide: true
output: html_notebook
md_extensions: +hard_line_breaks+pipe_tables
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    #, fig.align = "center"
    #, fig.width = 3.27, fig.height = 2.5, dev.args = list(pointsize = 10)
    #,cache = TRUE
    #, fig.width = 4.3, fig.height = 3.2, dev.args = list(pointsize = 10)
    #, fig.width = 6.3, fig.height = 6.2, dev.args = list(pointsize = 10)
    # works with html but causes problems with latex
    #,out.extra = 'style = "display:block; margin: auto"' 
    )
knitr::knit_hooks$set(spar = function(before, options, envir) {
    if (before) {
        par(las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
    }
})
```


# Background
## Random uncertainty
```{r echo=FALSE}
library(REddyProc)  
library(dplyr, quietly = TRUE)
EProc = local({load("StandardProcessing_fluxpart.RData"); get(ls()[1])})  
results = cbind(EProc$sExportData(), EProc$sExportResults())
```

Random uncertainty is the scatter of NEE for otherwise same conditions.
It includes measurement error and intrinsic stochasticity of the studied system.

There is low uncertainty during nighttime and higher uncertianty during daytime.
```{r}
plot( NEE_uStar_fsd ~ DateTime, slice(results,400:600))
```

## Uncertainty scales with flux
```{r}
plot( NEE_uStar_fsd ~ NEE_uStar_fall, slice(results, sample.int(nrow(results),400)))
```

This is called heteroscedasticy.

## Wrong aggregation without correlations
Error propagatin by adding variance results in an approximate reduction of the  mean (root mean square) 
$\sqrt{\bar{\sigma^2}}$ by $\sqrt{n}$.

```{r}
results %>% filter(NEE_uStar_fqc == 0) %>% summarise(
  nRec = sum(is.finite(NEE_uStar_f))
  , rms = sqrt(mean(NEE_uStar_fsd^2))
  , seMean = rms / sqrt(nRec)
  #, seMean = sqrt(sum(NEE_uStar_fsd^2))/nRec
  ) %>% select(nRec, rms, seMean)
```
Due to the large number of records, the estimated uncertainty is very low.


```{r}

```

