---
title: "Gebesee handson"
author: "Thomas Wutzler"
date: "`r Sys.Date()`"
output: html_notebook
---


# uStar Threshold estimation

## Preparing the data
First, the data is loaded. This example uses data that has been downloaded 
from http://www.europe-fluxdata.eu
and preprocessed by `fLoadEuroFlux16`, where the DateTime Column has been created, 
and the variables renamed to the BGC-convention (e.g. Tair instead of Ta).

```{r}
library(REddyProc)
data(DEGebExample)
summary(DEGebExample)
```

VPD was not given with the original dataset and is calculated from Tair and rH.
```{r}
DEGebExample$VPD <- fCalcVPDfromRHandTair(DEGebExample$rH, DEGebExample$Tair)
```

## Task 1a: Create the REddyProc class

named `EProc` with columns c('NEE','Rg','Tair','VPD', 'Ustar')

at Location LatDeg = 51.1, LongDeg = 10.9, 

and timezone one hour adead of GMT (TimeZoneHour = 1)
```{r}
?sEddyProc_initialize  # called by sEddyProc$new
?sEddyProc_sSetLocationInfo
### TO_COMPLETE
EProc <- sEddyProc$new('DE-Geb', DEGebExample, c('NEE','Rg','Tair','VPD', 'Ustar'))
EProc$sSetLocationInfo(LatDeg = 51.1, LongDeg = 10.9, TimeZoneHour = 1)  #Location of Gebesee
```

## Defining Seasons with different surface friction conditions 

The site is a crop site. The harvesting times are visible as sharp edges in the plots of NEE.

The micrometeorological conditions differ between the different cropping periods,
because the friction at the surface differs.
Also not the cropping periods do not correspond very well to seasons.
Hence, for the estimation of uStar-Thresholds, we apply a user-defined splitting 
of uStar-seasons. With function `usCreateSeasonFactorYdayYear` we provide the starting
points of the seasons.  
```{r}
plot( NEE ~ DateTime, DEGebExample )
```

## Task 1b: Define sesason

Base on the NEE plot define seasons with
starting at day 70, 210, 320 in 2004 and 70,180,320 in 2005 and 120,305 in 2006

Optional: Add horizontal lines to the NEE~DateTime plot.
```{r DEGeb_estUStar1a, spar = TRUE, message = FALSE, fig.width = 10}
?usCreateSeasonFactorYdayYear
### TO_COMPLETE
seasonStarts <- as.data.frame( do.call( rbind, list(
	  c(70,2004)
          ,c(210,2004)
          ,c(320,2004)
          ,c(70,2005)
          ,c(180,2005)
          ,c(320,2005)
          ,c(120,2006)
          ,c(305,2006) 		
)))
seasonFactor <- usCreateSeasonFactorYdayYear(
  DEGebExample$DateTime - 15*60, starts = seasonStarts)
# optional
seasonStartsDate <- fConvertTimeToPosix( data.frame(Year = seasonStarts[,2]
	, DoY = seasonStarts[,1], Hour = 0.25), 'YDH'
	, Year = "Year", Day = "DoY", Hour = "Hour")
plot( NEE ~ DateTime, DEGebExample )
abline( v = seasonStartsDate$DateTime)
```

## Task 1c: Estimate the distribution of u* thresholds

Specify the seasons defined before.

For saving time, here use only 30 bootstrap samples (argument nSample) and 
estimate only the 10th and 90th percentile (argument probs)

Print the estimated thresholds.


```{r DEGeb_estUStar1b}
?sEddyProc_sEstimateUstarScenarios
?sEddyProc_sGetEstimatedUstarThresholdDistribution
### TO_COMPLETE
EProc$sEstimateUstarScenarios( 
  seasonFactor = seasonFactor, nSample = 30L, probs = c(0.1,0.9))
EProc$sGetEstimatedUstarThresholdDistribution()
```

# Gapfilling

## Task 2a: Use seasonal u* thresholds

- display the default annually used thresholds
- specify using seasonsal thresholds
- display the scenarios again


```{r, message = FALSE}
?sEddyProc_sGetUstarScenarios
?sEddyProc_useSeaonsalUStarThresholds
### TO_COMPLETE
EProc$sGetUstarScenarios()
EProc$useSeaonsalUStarThresholds()
EProc$sGetUstarScenarios()
```

## Task 2b: Perform gapfilling for NEE

Estimate random error also for non-gap records.
```{r, message=FALSE}
?sEddyProc_sMDSGapFillUStarScens
### TO_COMPLETE
EProc$sMDSGapFillUStarScens("NEE", FillAll = TRUE)
```

What columns have been created?
```{r}
colnames(EProc$sExportResults())
```

For an explanation of the column names see [output format description](https://www.bgc-jena.mpg.de/bgi/index.php/Services/REddyProcWebOutput)
- NEE_<scenario>_f: gaps replaced by modeled values (gapfilled)
- NEE_<scenario>_fall: all NEE replaced by modeled values
- NEE_<scenario>_fqc: quality flag: 0 observations, 1 good quality of gapfilling 

## Task 2c: Produce a fingerprint plot of gapfilled values

First a single plot in document
- for 90% quantile $u_{*Th}$ (suffix U90)
- for year 2005

```{r}
?sEddyProc_sPlotFingerprintY
### TO_COMPLETE
EProc$sPlotFingerprintY('NEE_U90_f', Year = 2005)
```

Next, produce pdf-files with legend for all years in subdirectory "plotsHandsOn"
```{r}
?sEddyProc_sPlotFingerprint
### TO_COMPLETE
EProc$sPlotFingerprint('NEE_U90_f', Dir = "plotsHandsOn")
```

# Flux partitioning

## Task 3a: Prepare for partitioning
Specify the Location and time zone (51.1N, 10.0W, 1hour ahead of GMT)

Gapfill the necessary meteorological input variables (Rg, Tair, VPD). Here
we do not need computing the uncertainty of the non-filled records.

```{r, message=FALSE}
?sEddyProc_sSetLocationInfo
?sEddyProc_sMDSGapFill  # note the FillAll argument 
### TO_COMPLETE
EProc$sSetLocationInfo(LatDeg = 51.1, LongDeg = 10.9, TimeZoneHour = 1)  #Location of Gebesee
EProc$sMDSGapFill('Rg', FillAll = FALSE)     
EProc$sMDSGapFill('Tair', FillAll = FALSE)     
EProc$sMDSGapFill('VPD', FillAll = FALSE)     
```

## Task 3b: Nighttime partitioning
Perform the nighttime partitioning

```{r, message=FALSE}
?sEddyProc_sMRFluxPartitionUStarScens
### TO_COMPLETE
EProc$sMRFluxPartitionUStarScens()
```

## Task 3c: Plotting the head GPP
Query the result data.frame to variable dsResults
```{r}
?sEddyProc_sExportResults
### TO_COMPLETE
dsResults <- EProc$sExportResults()
```

Plot the head of GPP for U90 scenario against time (DEGebExample$DateTime)
```{r}
nRec = 48*4 # 4 days of half-hours
plot(head(dsResults$GPP_U90_f,nRec) ~ head(DEGebExample$DateTime, nRec), type = "l")
```


## Task3d: Daytime partitioning
Perform the daytime partitioning.

```{r, message=FALSE}
?sEddyProc_sGLFluxPartitionUStarScens
### TO_COMPLETE
EProc$sGLFluxPartitionUStarScens()
```

Bonus Task: Repeat with fixed Temperature Senstivity to 80 $\pm$40 K
```{r eval=FALSE}
?partGLControl
### TO_COMPLETE
EProc$sGLFluxPartitionUStarScens(
  controlGLPart = partGLControl(
    fixedTempSens = data.frame(E0 = 80, sdE0 = 40, RRef = NA_real_))
  ,isWarnReplaceColumns = FALSE
)
```

## Task 3e: Produce fingerprint plots of GPP_DT and Reco_DT 
- For the original non-bootstrapped data uStar scenario (suffix uStar)
- In Sub-Directory `plotsHandsOn`
```{r}
?sEddyProc_sPlotFingerprint
### TO_COMPLETE
EProc$sPlotFingerprint("GPP_DT_uStar",Dir = "plotsHandsOn")
EProc$sPlotFingerprint("Reco_DT_uStar",Dir = "plotsHandsOn")
```

## Task 3f: Save the results together with the data
as tab-delimted text file "plotsHandsOn/DEGeb_Part.txt"
Store the combined data and results in variable 'results'
```{r}
?sEddyProc_sExportResults
?sEddyProc_sExportData
?fWriteDataframeToFile
### TO_COMPLETE
dsRes <- EProc$sExportResults()
dsData <- EProc$sExportData()
results <- cbind(dsData, dsRes)
fWriteDataframeToFile(results, "DE-Geb_Part.txt", Dir = "plotsHandsOn")
```

```{r include=FALSE}
#if (file.exists("plotsHandsOn/DE-Geb_Part.txt"))
#  cat(readLines("plotsHandsOn/DE-Geb_Part.txt", n = 5))
```
```{r include=FALSE, eval=FALSE}
# execute only interactively: save class to have an restore point
save(EProc, file = "plotsHandsOn/DE-Geb_EProcPart.RData")
EProc = local({load("plotsHandsOn/DE-Geb_EProcPart.RData"); get(ls()[1])})  
```

# Bias with $u_{*Th}$
## Focus on year 2005
```{r}
results$year <- as.POSIXlt(results$DateTime)$year + 1900
res05 <- subset(results, year == 2005)
```

## Task 4a: compute the annual mean NEE for each $u_{*Th}$ scenario

For the mean we can use gap-filled series in column `NEE_<suffix>_f`
```{r}
scens = c("uStar","U10","U90")
NEE_UStar <- 
### TO_COMPLETE  
  sapply(scens, function(suffix){
  colName = paste0("NEE_",suffix,"_f")
  mean(res05[[colName]])
})
NEE_UStar
```

## Task 4b: Compute statistic across the obtained aggregates
median, standard deviation and relative error
```{r}
### TO_COMPLETE
c(mean(NEE_UStar), sd(NEE_UStar), sd(NEE_UStar)/abs(mean(NEE_UStar)))
```


# Random uncertainty aggregation

## Task 5a: approximate the error terms in results 
For each good measurement, i.e. where `NEE_uStar_fqc == 0` we have
measured NEE that includes random fluctuation `NEE_uStar_orig` 
and gap-filling estimate `NEE_uStar_fall`
of the expected flux for the conditions at this time.

Create a new column `resid` that stores the difference for good conditions
and NA for other records.

```{r}
n <- sum(results$NEE_uStar_fqc == 0)  # number of good records
?ifelse
### TO_COMPLETE
results$resid = ifelse(
  results$NEE_uStar_fqc == 0, results$NEE_uStar_orig - results$NEE_uStar_fall, NA )
```

## Task 5b: compute the empirical autocorrelation function
```{r}
library(lognorm)
?computeEffectiveAutoCorr
### TO_COMPLETE
rho = computeEffectiveAutoCorr(results$resid)
plot(rho[-1])
```
## Task 5c: Compute the effective number of effice observation
To variable `nEff`

and compare to the number of good observations `n`.

```{r}
?computeEffectiveNumObs
### TO_COMPLETE
nEff <- computeEffectiveNumObs(results$resid, na.rm = TRUE)
c(nEff, n)
```

## Task 5d: Compute its effective number of observations for year 2005
Use the autocorrelation function that you determined on the 
entire dataset before (`rho`)
```{r}
?computeEffectiveNumObs # note argument effAcf
### TO_COMPLETE
nEff <- computeEffectiveNumObs(res05$resid, na.rm = TRUE, effAcf = rho)
```

## Task 5e Compute the mean annual NEE and its standard deviation for 2005
For scenario $u_{*Th}$ threshold scenario `uStar`
Report also the relative error, i.e. coefficient of variation.

Consider gap-filled values in the mean. 
But make sure to not use gap-filled records in unertainty estimation.

Remember that the root-mean-squared averaged standard deviation decreases 
by factor $\sqrt{nEff -1}$.

```{r}
sdGood = res05$NEE_uStar_fsd[res05$NEE_uStar_fqc == 0]
### TO_COMPLETE
NEE = mean(res05$NEE_uStar_f)
sdNEE = sqrt(mean(sdGood^2)) / sqrt(nEff - 1)
c(mean = NEE, sd = sdNEE, cv = sdNEE/abs(NEE))
```

# Combined uncertainty
## Task 6a Compute the combined uncertaint of random and $u_{*Th}$

Remember that standard deviation of independent variables adds in squares.
```{r}
sdNEEUStar = sd(NEE_UStar)
sdNEECombined <-
### TO_COMPLETE
   sqrt(sdNEEUStar^2 + sdNEE^2)
sdNEECombined 
```



