---
title: "Gebesee handson"
author: "Thomas Wutzler"
date: "`r Sys.Date()`"
output: html_notebook
---


# uStar Threshold estimation

## Preparing the data
First, the data is loaded. This example uses data that has been downloaded 
from http://www.europe-fluxdata.eu
and preprocessed by `fLoadEuroFlux16`, where the DateTime Column has been created, 
and the variables renamed to the BGC-convention (e.g. Tair instead of Ta).

```{r}
library(REddyProc)
data(DEGebExample)
summary(DEGebExample)
```

VPD was not given with the original dataset and is calculated from Tair and rH.
```{r}
DEGebExample$VPD <- fCalcVPDfromRHandTair(DEGebExample$rH, DEGebExample$Tair)
```

## Task 1a: Create the REddyProc class

named `EProc` with columns c('NEE','Rg','Tair','VPD', 'Ustar')

at Location LatDeg = 51.1, LongDeg = 10.9, 

and timezone one hour adead of GMT (TimeZoneHour = 1)
```{r}
?sEddyProc_initialize  # called by sEddyProc$new
?sEddyProc_sSetLocationInfo
### TO_COMPLETE
EProc <- sEddyProc$new('DE-Geb', DEGebExample, c('NEE','Rg','Tair','VPD', 'Ustar'))
EProc$sSetLocationInfo(LatDeg = 51.1, LongDeg = 10.9, TimeZoneHour = 1)  #Location of Gebesee
```

## Defining Seasons with different surface friction conditions 

The site is a crop site. The harvesting times are visible as sharp edges in the plots of NEE.

The micrometeorological conditions differ between the different cropping periods,
because the friction at the surface differs.
Also not the cropping periods do not correspond very well to seasons.
Hence, for the estimation of uStar-Thresholds, we apply a user-defined splitting 
of uStar-seasons. With function `usCreateSeasonFactorYdayYear` we provide the starting
points of the seasons.  
```{r}
plot( NEE ~ DateTime, DEGebExample )
```

## Task 1b: Define sesason

Base on the NEE plot define seasons with
starting at day 70, 210, 320 in 2004 and 70,180,320 in 2005 and 120,305 in 2006

Optional: Add horizontal lines to the NEE~DateTime plot.
```{r DEGeb_estUStar1a, spar = TRUE, message = FALSE, fig.width = 10}
?usCreateSeasonFactorYdayYear
### TO_COMPLETE
seasonStarts <- as.data.frame( do.call( rbind, list(
	  c(70,2004)
          ,c(210,2004)
          ,c(320,2004)
          ,c(70,2005)
          ,c(180,2005)
          ,c(320,2005)
          ,c(120,2006)
          ,c(305,2006) 		
)))
seasonFactor <- usCreateSeasonFactorYdayYear(
  DEGebExample$DateTime - 15*60, starts = seasonStarts)
# optional
seasonStartsDate <- fConvertTimeToPosix( data.frame(Year = seasonStarts[,2]
	, DoY = seasonStarts[,1], Hour = 0.25), 'YDH'
	, Year = "Year", Day = "DoY", Hour = "Hour")
plot( NEE ~ DateTime, DEGebExample )
abline( v = seasonStartsDate$DateTime)
```

## Task 1c: Estimate the distribution of u* thresholds

Specify the seasons defined before.

For saving time, here use only 30 bootstrap samples (argument nSample) and 
estimate only the 10th and 90th percentile (argument probs)

Print the estimated thresholds.


```{r DEGeb_estUStar1b}
?sEddyProc_sEstimateUstarScenarios
?sEddyProc_sGetEstimatedUstarThresholdDistribution
### TO_COMPLETE
EProc$sEstimateUstarScenarios( 
  seasonFactor = seasonFactor, nSample = 30L, probs = c(0.1,0.9))
EProc$sGetEstimatedUstarThresholdDistribution()
```

# Gapfilling

## Task 2a: Use seasonal u* thresholds

- display the default annually used thresholds
- specify using seasonsal thresholds
- display the scenarios again


```{r, message = FALSE}
?sEddyProc_sGetUstarScenarios
?sEddyProc_useSeaonsalUStarThresholds
### TO_COMPLETE
EProc$sGetUstarScenarios()
EProc$useSeaonsalUStarThresholds()
EProc$sGetUstarScenarios()
```

## Task 2b: Perform gapfilling for NEE

Estimate random error also for non-gap records.
```{r, message=FALSE}
?sEddyProc_sMDSGapFillUStarScens
### TO_COMPLETE
EProc$sMDSGapFillUStarScens("NEE", FillAll = TRUE)
```

## Task 2c: Produce a fingerprint plot of gapfilled values

First a single plot in document
- for 90% quantile $u_{*Th}$ (suffix U90)
- for year 2005

```{r}
?sEddyProc_sPlotFingerprintY
### TO_COMPLETE
EProc$sPlotFingerprintY('NEE_U90_f', Year = 2005)
```

Next, produce pdf-files with legend for all years in subdirectory "plotsHandsOn"
```{r}
?sEddyProc_sPlotFingerprint
### TO_COMPLETE
EProc$sPlotFingerprint('NEE_U90_f', Dir = "plotsHandsOn")
```

# Flux partitioning

## Task 3a: Prepare for partitioning
Specify the Location and time zone (51.1N, 10.0W, 1hour ahead of GMT)

Gapfill the necessary meteorological input variables (Rg, Tair, VPD). Here
we do not need computing the uncertainty of the non-filled records.

```{r, message=FALSE}
?sEddyProc_sSetLocationInfo
?sEddyProc_sMDSGapFill  # note the FillAll argument 
### TO_COMPLETE
EProc$sSetLocationInfo(LatDeg = 51.1, LongDeg = 10.9, TimeZoneHour = 1)  #Location of Gebesee
EProc$sMDSGapFill('Rg', FillAll = FALSE)     
EProc$sMDSGapFill('Tair', FillAll = FALSE)     
EProc$sMDSGapFill('VPD', FillAll = FALSE)     
```

## Task 3b: Nighttime partitioning
Perform the nighttime partitioning

```{r, message=FALSE}
?sEddyProc_sMRFluxPartitionUStarScens
### TO_COMPLETE
EProc$sMRFluxPartitionUStarScens()
```

## Task 3c: Plotting the head GPP
Query the result data.frame to variable dsResults
```{r}
?sEddyProc_sExportResults
### TO_COMPLETE
dsResults <- EProc$sExportResults()
```

Plot the head of GPP for U90 scenario against time (DEGebExample$DateTime)
```{r}
nRec = 48*4 # 4 days of half-hours
plot(head(dsResults$GPP_U90_f,nRec) ~ head(DEGebExample$DateTime, nRec), type = "l")
```


## Task3d: Daytime partitioning
Perform the daytime partitioning.

```{r, message=FALSE}
?sEddyProc_sGLFluxPartitionUStarScens
### TO_COMPLETE
EProc$sGLFluxPartitionUStarScens()
```

Bonus Task: Repeat with fixed Temperature Senstivity to 80 $\pm$40 K
```{r eval=FALSE}
?partGLControl
### TO_COMPLETE
EProc$sGLFluxPartitionUStarScens(
  controlGLPart = partGLControl(
    fixedTempSens = data.frame(E0 = 80, sdE0 = 40, RRef = NA_real_))
  ,isWarnReplaceColumns = FALSE
)
```

## Task 3e: Produce fingerprint plots of GPP_DT and Reco_DT 
- For the original non-bootstrapped data uStar scenario (suffix uStar)
- In Sub-Directory `plotsHandsOn`
```{r}
?sEddyProc_sPlotFingerprint
### TO_COMPLETE
EProc$sPlotFingerprint("GPP_DT_uStar",Dir = "plotsHandsOn")
EProc$sPlotFingerprint("Reco_DT_uStar",Dir = "plotsHandsOn")
```

## Task 3f: Save the results together with the data
as tab-delimted text file "plotsHandsOn/DEGeb_Part.txt"
```{r}
?sEddyProc_sExportResults
?sEddyProc_sExportData
?fWriteDataframeToFile
### TO_COMPLETE
dsRes <- EProc$sExportResults()
dsData <- EProc$sExportData()
fWriteDataframeToFile(cbind(dsData, dsRes), "DE-Geb_Part.txt", Dir = "plotsHandsOn")
```

```{r include=FALSE}
#if (file.exists("plotsHandsOn/DE-Geb_Part.txt"))
#  cat(readLines("plotsHandsOn/DE-Geb_Part.txt", n = 5))
```

# Uncertainty aggregation

# Before
After setting the uStar scenarios, they will be used by method 
`sMDSGapFillUStarScens`.
```{r DEGeb_gapFillBoot1, cache = TRUE, message = FALSE, output = 'hide'}
EProc$sMDSGapFillUStarScens('NEE', FillAll = FALSE)
```

Additional output columns are produced for each uStar quantile. 

```{r}
grep("^NEE.*_f$", colnames( EProc$sExportResults()), value = TRUE )
```		

Several methods provide processing steps for all scenarios.
In addition, method `sApplyUStarScen` calls a user-specified function repeatedly
with a modified argument `suffix`. 
For example, the flux partitioning across all threshold scenarios can be
invoked by the following code to produce several GPP estimates.

```{r DEGeb_fluxPart1, cache = TRUE, message = FALSE}
EProc$sMDSGapFill('Tair', FillAll = FALSE)
EProc$sApplyUStarScen( EProc$sMRFluxPartition )
#grep("U10", colnames(EProc$sExportResults()), value = TRUE) 	
grep("^GPP.*_f$", colnames( EProc$sExportResults()), value = TRUE )
```

Using change point detection instead of moving point method for UStar Threshold estimation
------------------------------------------------------------------------------------------

The package also provides another method of estimating the point where NEE saturates 
with increasing uStar.
With the ChangePointDetection (CPT) method, the data is not binned by classes of 
uStar but the changepoint
is estimated based on the entire subset within one seasons and one temperature class.
The user invokes this method by specifying argument 
`ctrlUstarEst.l = usControlUstarEst(isUsingCPTSeveralT = TRUE)`
to `EstUstarThold` or `sEstUstarThresholdDistribution`.

The CPT method is usually yields higher thresholds and marks more data as Gap. 
  
```{r DEGeb_estUStarCPT, cache = TRUE}
EProc <- sEddyProc$new(
  'DE-Geb', DEGebExample, c('NEE','Rg','Tair','VPD', 'Ustar'))
resUStar <- EProc$sEstUstarThold(
					ctrlUstarEst = usControlUstarEst(isUsingCPTSeveralT = TRUE)
					, seasonFactor = seasonFactor
			)
#(uStarThCP <- usGetSeasonalSeasonUStarMap(resUStar))
EProc$useSeaonsalUStarThresholds()
EProc$sGetUstarScenarios()
```


		




