---
title: "Uncertainty in annual aggregates"
author: "Thomas Wutzler"
output:
  ioslides_presentation:
    css: slidesTw.css
    wide: true
#output: html_notebook
md_extensions: +hard_line_breaks+pipe_tables
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    #, fig.align = "center"
    #, fig.width = 3.27, fig.height = 2.5, dev.args = list(pointsize = 10)
    #,cache = TRUE
    #, fig.width = 4.3, fig.height = 3.2, dev.args = list(pointsize = 10)
    #, fig.width = 6.3, fig.height = 6.2, dev.args = list(pointsize = 10)
    # works with html but causes problems with latex
    #,out.extra = 'style = "display:block; margin: auto"' 
    )
knitr::knit_hooks$set(spar = function(before, options, envir) {
    if (before) {
        par(las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
    }
})
```

# Bias with $u_{*Th}$

## Motivation
```{r echo=FALSE, message=FALSE}
library(REddyProc)  
library(dplyr, quietly = TRUE)
EProc = local({load("StandardProcessing_fluxpart.RData"); get(ls()[1])})  
results = cbind(EProc$sExportData(), EProc$sExportResults())
```
We get a different annually aggregated NEE for each quantile of $u_{*Th}$.
Usually NEE is lower with lower threshold.

For an uncertainty estimate we must compute the annual aggregate for
each of the scenarios.

Getting the scenario names:
```{r, results="hold"}
(uStarScens = EProc$sGetUstarSuffixes())
```

## Aggregate Reco for each scenario {.smaller}
```{r}
fMean <- function(suffix){
  colName = paste0("Reco_",suffix)
  #colName = paste0("GPP_",suffix,"_f")
  mean(results[[colName]])
}
#fMean("U50")
meansUStar <- unlist(EProc$sApplyUStarScen(fMean))
names(meansUStar) <- uStarScens
meansUStar
```

```{r}
c(median(meansUStar), sd(meansUStar), 100*sd(meansUStar)/abs(median(meansUStar)))
```

Here only small relative error of 1.5% in respiration introduced by 
unknown $u_{*Th}$

## More robust analysis of $u_{*Th}$ effects
For a thorough analysis, we would have needed 
to specify more than 3 quantiles 
when we estimated the $u_{*Th}$ distribution.

But this would have multiplied the computation time.

```{r eval=FALSE}
EProc$sEstimateUstarScenarios(
    nSample = 200L, probs = seq(0.125,0.975, length.out = 30))
```

# Random uncertainty
## Random uncertainty {.smaller}
Random uncertainty is the scatter of NEE for otherwise same conditions.
It includes measurement error and intrinsic stochasticity of the studied system.

There is low uncertainty during nighttime and higher uncertianty during daytime.
```{r}
plot( NEE_uStar_fsd ~ DateTime, slice(results,400:600))
```

## Uncertainty scales with flux (heteroscadastic)
```{r}
plot( NEE_uStar_fsd ~ NEE_uStar_fall, slice(results, sample.int(nrow(results),400)))
```

## Wrong aggregation without correlations 
Error propagatin by adding variance results in an approximate reduction of the  mean (root mean square) 
$\sqrt{\bar{\sigma^2}}$ by $\sqrt{n}$.

```{r}
stat1 <- results %>% filter(NEE_uStar_fqc == 0) %>% summarise(
  nRec = sum(is.finite(NEE_uStar_f))
  , rms = sqrt(mean(NEE_uStar_fsd^2, na.rm = TRUE))
  , seMean = rms / sqrt(nRec)
  #, seMean = sqrt(sum(NEE_uStar_fsd^2))/nRec
  ) %>% select(nRec, seMean) %>% unlist()
stat1["seMean"]
```
Due to the large number of records, the estimated uncertainty is very low.


## Considering autocorrelation {.smaller}
Errors of subsequent measurements tend to have the same direction and magnitude.
Subsequent errors tend to be similar.

We approximate the error terms by the difference between original measurement
and value from Marginal Distribution samples during gapfilling.

```{r}
results <- results %>% 
  mutate(resid = ifelse(NEE_uStar_fqc == 0, NEE_uStar_orig - NEE_uStar_fall, NA ))
```
```{r echo=FALSE, fig.height=3}
plot(resid ~ DateTime, slice(results, 300:350), type = "l")
abline(h = 0, col = "grey")
```

## Correlation between subsequent errors
```{r}
n <- nrow(results)
plot(NEE_uStar_fsd[-n] ~ NEE_uStar_fsd[-1], results); abline(a = 0,b = 1)
```

## Autocorrelation function
They are not independent of each other. The strenght of their dependence 
can be expressed by values $\rho_i$ 
$$
x_{i} = \rho_1 x_{i-1} + \rho_2 x_{i-2} + ... + \epsilon_i
$$

The series $\rho$ is called the autocorrelation function.

## Empirical autocorrelation function {.smaller}

The series of estimates of those coefficients from the data is called the 
empirical autocorrelation function.
But only the first terms of the empricial functions are reliable.
(Take only the ones before the first negative coefficient)

```{r fig.height=3}
#rho = acf(results$resid, lag.max = 20, main = "", na.action = na.exclude)
library(lognorm)
rho = computeEffectiveAutoCorr(results$resid)
plot(rho[-1])
```

## Error propagation
If the observations are not independent the average of the mean 
does not scale by $\sqrt{n}$ but by approximately 
$\sqrt{n_{eff}}$ with $n_{eff} < n$.

$$
\sigma(m) = \frac{ \sqrt{\bar{\sigma^2}} }{ \sqrt{n_{eff} -1}}
$$

## Effectively fewer observations
```{r}
library(lognorm)
nEff <- computeEffectiveNumObs(results$resid, na.rm = TRUE)
c( nEff = nEff, nObs = sum(is.finite(results$resid)))
```
Due to the high autocorrelation, the entire year of NEE records have only
fewer effective observations.

When computing correlations or effective number of observations, make sure
to use a complete time series with a records for each equidistant time step.
Missing values will 

## Larger aggregated error
```{r}
stat2 <- results %>% filter(NEE_uStar_fqc == 0) %>% summarise(
  rms = sqrt(mean(NEE_uStar_fsd^2, na.rm = TRUE))
  , seMeanCor = rms / sqrt(nEff - 1)
  ) %>% select(seMeanCor) %>% unlist()
NEE <- mean(results$NEE_uStar_f)
c(stat2["seMeanCor"], stat1["seMean"], cv = stat2["seMeanCor"]/abs(NEE))
```

# Combined uncertainty

## Add in squares

The uncertainty due to random fluctuations can be assumed independent
to the uncertainty introduced by the uncertain $u_{*Th}$. 
Then error propgation can be done by adding variances.

$$
\sigma_{Combined}(x) = \sqrt{\sigma^2_{u_*}(x) + \sigma^2_{Random}(x)}
$$

```{r}
sdNEERandom <- sd(unlist(EProc$sApplyUStarScen(function(suffix){
  mean(results[[paste0("NEE_",suffix,"_f")]])
})))
sdNEE <- sqrt(stat2["seMeanCor"]^2 + sdNEERandom^2)
sdNEE
```



